---
title: "naglu 9 month behaviour"
author: "Karissa Barthelson"
date: "2021-10-26"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## setup
```{r}
library(tidyverse)
library(magrittr)
library(readxl)
library(here)
library(scales)
library(ggpubr)
library(ggeasy)
library(ggrepel)
library(ggfortify)
library(ggbeeswarm)
library(ggforce)
library(kableExtra)

# stat analysis
library(broom)
library(lme4)
library(performance)
library(car)
library(emmeans)
library(glmmTMB)
library(MASS)

# set the default theme for ggplot as theme_bw
theme_set(theme_bw())

# read in Ewans metadata speadsheet. make all the metadata as factors as well
meta <- 
  read_excel("data/metadata/naglu 9 month behavioural metadata.xlsx") %>% 
  mutate(
    fish_id = as.factor(fish_id), 
    Day = as.factor(Day), 
    Time = as.factor(Time), 
    Tank = as.factor(Tank), 
    Sex = as.factor(Sex), 
    Genotype = factor(Genotype, levels = c("wt", "het", "hom"))
  )
# prepare the final data object which contains genotype
final_data <- read_csv("data/processed_data/naglu/9m/final_output.csv") %>% 
  dplyr::select(-1) %>% 
  mutate(fish_id = factor(fish_id)) %>% 
  left_join(meta) %>% 
  dplyr::filter(Keep == "Y") 

# make an object which converts the final data to long format. for easier plotting in ggplot
final_data_long <- final_data %>%
  gather(key = "tetras", value = "Count", # convert to long format
         grep("[L|R]{4}", 
              colnames(.))
         )

# also make an object which sums the tetragrams over the hour
final_data_summedoverbins <- 
  final_data %>%
  gather(key = "tetras", value = "Count", # convert to long format
         grep("[L|R]{4}", # select the columns which contain a L or a R four times
              colnames(.))
  ) %>% 
  group_by(fish_id, tetras) %>% 
  mutate(x = sum(Count)) %>% # sum the tetragram counts per fish_id
  dplyr::select(colnames(meta), tetras, x) %>% 
  ungroup() %>% 
  dplyr::distinct(fish_id, tetras, .keep_all = T)
```

# Introduction

In this analysis, I will analyse Ewans raw data generated for zebrafish arising from a lay from the same *naglu* A603fs het x het parents in the Zantiks Y-maze. 

Briefly, zebrafish are raised from this cross together in the same tank, across `r meta$Tank %>% unique %>% length` lays. Each lay has come from the same parents are approx. 1 week apart in age. Each tank contained fish either WT, heterozygous or homozgous for the mutation in a 1:2:1 ratio. 

Fish were isolated for 30 mins two at a time, then placed in the maze for 1 hour. Raw data was collected, and metadata was recorded (i.e. sex, time, day). These raw data spreadsheets were batch processed in the naglu 3, BatchProcess R script to produce the `final_data` object. This object contains the tetragram frequencies (see below). Note that fish were genotyped after data collection so that we are blinded from any observer bias until after raw data collection. Addiotianlly, the videos were manually observed to ensure proper tracking. Any fish where the tracking was not perfect was omitted. 

```{r}
list.files(path = "data/raw_data/naglu/9m/", 
           pattern = "*.csv", 
           recursive = TRUE, full.names = TRUE) %>% 
  .[1] %>% 
  read_csv(skip =4, col_names = F) %>% 
  head %>% 
  dplyr::select(-X2, -X3) %>% 
  set_colnames(c("Time", "fish",'enter or exit', "zone")) %>% 
  kable(caption = "example of raw data") %>% 
  kable_styling(full_width = F)
```


```{r}
final_data %>% 
  head %>% 
  kable(caption = "Example of the processed data") %>% 
  kable_styling() 
```

# check genotype proportions

First, I will check that genotype proportions are as expected. A few less homozygous fish are observed than expected. However, this is partly due to that they were omitted because the software did not track properly. 

```{r}
meta %>% 
  ggplot(aes(Genotype)) +
  geom_bar(colour = "black", 
           aes(fill = Tank)) +
  scale_fill_viridis_d() +
  ggtitle("Total number of fish per genotype in the study") 
```

# Visualisation of raw data

Cleal et al. showed that zebrafish naturally perform more of the alternation tetragrams (LRLR and RLRL) in a Y-maze. This is also obsevred in our data here. A few fish may possibly be outliers. 

```{r}
final_data_summedoverbins %>%
  ggplot(aes(x = tetras, y = x)) +
  geom_jitter(aes(colour = tetras,
                  shape = Sex)) +
  geom_boxplot(outlier.shape = NA,
               # aes(fill = Genotype)
               fill = NA
  ) +
  geom_label_repel(aes(label = fish_id), # only label the points if the number of tetragrams was over 200
                   data = . %>% 
                     dplyr::filter(x > 200)) +
scale_fill_viridis_d() +
  labs(y = "Number of tetragrams", 
       colour = "Tetragram",
       x = "Tetragram")+
  theme(legend.position = "bottom")  +
  ggtitle("Total number of 16 possible tetragrams performed by zebrafish in a Y-maze\nduring a 1 hour search period") 
```


We can also overlay the genotype boxplots as shown below. Looks like the heterozyous and homozygous fish perform more alternations than the WT fish. I will look into this more later in the analysis 

```{r}
final_data_summedoverbins %>%
  ggplot(aes(x = tetras, y = x)) +
  geom_jitter(aes(colour = tetras,
                  shape = Sex)) +
  geom_boxplot(outlier.shape = NA,
               aes(fill = Genotype) # fill the boxplots by genotype
  ) +
  geom_label_repel(aes(label = fish_id), # only label the points if the number of tetragrams was over 200
                   data = . %>% 
                     dplyr::filter(x > 200)) +
  scale_fill_viridis_d() +
  labs(y = "Number of tetragrams",
       colour = "Tetragram",
       x = "Tetragram")+
  theme(legend.position = "bottom")  +
  ggtitle("Total number of 16 possible tetragrams performed by zebrafish in a Y-maze\nduring a 1 hour search period") 
```

The alternation tetragrams are the tetragram of interest (the measure of working memory). Below indicates the number of tetragrams performed by zebrafish across the 6 x 10 min blocks of the hour they spent in the maze.

```{r}
final_data_long %>%
  dplyr::distinct(fish_id, bin, .keep_all = T) %>% 
  ggplot(aes(x = Genotype, y = alts, fill = Genotype)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(shape = Sex)) +
  geom_label_repel(aes(label = fish_id), # only label the points if the number of alts was over 100
                   fill = NA, 
                   data = . %>%
                     dplyr::filter(alts > 100)) +
  facet_wrap(~bin, nrow = 1) +
  theme(axis.text.x  = element_text(angle = 45, hjust = 1, vjust = 1)) +
  ylab("Frequency of alternation tetragrams (LRLR + RLRL)") +
  scale_fill_viridis_d()
```

We can also display this as the *relative* amount of alternation tetragrams performed by zebrafish in a 1 hour search period. This will control for how active each fish is.

```{r}
final_data_long %>%
  dplyr::distinct(fish_id, bin, .keep_all = T) %>% 
  ggplot(aes(x = Genotype, y = rel_alts, fill = Genotype)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(shape = Sex)) +
  facet_wrap(~bin, nrow = 1) +
  ylab("Relative alteration frequency (LRLR + RLRL)") +
  theme(axis.text.x  = element_text(angle = 45, hjust = 1, vjust = 1))  +
  scale_fill_viridis_d()
```

# Locomotor defect

Alteration to locomotor activity could be a confounding effect with genotype to alternation (LRLR + RLRL) frequency. If the mutant fish had some form of locomotor change, they could be performing more or less alternation tetragrams not due to impairment of spatial working memory. To account for this, I will test whether the genotype of the fish had a significant effect on 1) the total number of turns the fish performed in the hour spent in the maze or 2) the average time spent in each zone of the maze.

## total number of turns
```{r}
final_data_summedoverbins %>%
  group_by(fish_id) %>%
  mutate(total_turns = sum(x)) %>%
  dplyr::select(colnames(meta), total_turns) %>%
  unique %>% 
  ggplot(aes(x = Genotype,y = total_turns)) +
  geom_violin(aes(fill = Genotype),
               alpha = 0.5) +
  geom_boxplot(aes(colour = Genotype),
               fill = NA, outlier.shape = NA,
               width = 0.25,
               colour= "black") +
  geom_quasirandom(aes(shape = Sex)) +
  scale_fill_viridis_d(option = "viridis") +
  scale_colour_viridis_d(option = "viridis") +
  ylab("Total number of turns") +
  ggtitle("Total number of turns performed by fish in the Y-maze in 1 hour")

final_data %>% 
  ggplot(aes(x = Genotype, y = total_turns)) +
  geom_violin(aes(fill = Genotype),
              alpha = 0.5) +
  geom_boxplot(aes(colour = Genotype),
               fill = NA,
               width = 0.25,
               outlier.shape = NA,
               colour= "black") +
  geom_quasirandom(aes(shape = Sex)) +
  geom_label_repel(aes(label = fish_id), 
                   data = . %>% 
                     dplyr::filter(total_turns > 300)) +
  facet_wrap(~bin, nrow = 1) +
  scale_fill_viridis_d(option = "viridis") +
  scale_colour_viridis_d(option = "viridis") +
  ylab("Total number of turns") +
  ggtitle("Total number of turns performed by fish in the Y-maze in 1 hour")

final_data %>% 
  ggplot(aes(x = Sex, y = total_turns)) +
  geom_violin(aes(fill = Sex),
              alpha = 0.5) +
  geom_boxplot(aes(colour = Sex),
               fill = NA,
               width = 0.25,
               outlier.shape = NA,
               colour= "black") +
  geom_quasirandom(aes(shape = Sex)) +
  geom_label_repel(aes(label = fish_id), 
                   data = . %>% 
                     dplyr::filter(total_turns > 300)) +
  facet_wrap(~bin, nrow = 1) +
  scale_fill_viridis_d(option = "viridis") +
  scale_colour_viridis_d(option = "viridis") +
  ylab("Total number of turns") +
  ggtitle("Total number of turns performed by fish in the Y-maze in 1 houur")

final_data_summedoverbins %>% 
  group_by(fish_id) %>%
  mutate(total_turns = sum(x)) %>%
  dplyr::select(colnames(meta), total_turns) %>%
  unique %>% 
  glm.nb(total_turns ~ Genotype + Sex + Tank,
         data = .) %>%
  summary %>%
  .$coef %>%
  kable(caption = "Negative binomial GLM model coefficients overall in one hour" ) %>%
  kable_styling() %>% 
  row_spec(row = c(2,3), bold = T, color = "black", background = "lightblue")

final_data %>% 
  glm.nb(total_turns ~ (bin + Genotype)^2 + Sex + Tank,
         data = .) %>%
  summary %>%
  .$coef %>%
  kable(caption = "Negative binomial GLM model coefficients. accounting for differences across the hour" ) %>%
  kable_styling() %>% 
  row_spec(row = c(3,4,9,10), bold = T, color = "black", background = "lightblue")
```

## time spent in each zone of the Y-maze

From inspection of the **total** and **average** time spent in each zone, something may have gone wrong with fish 56, or it is a non-responder. I therefore tried to inspect the video to have a look at what is going on. But it seems the video is corrupted and cant be viewed. It appears that this fish may have spent a bit of time sitting in one of the arms of the maze. 

```{r}
read_csv("data/processed_data/naglu/9m/time_in_zone.csv") %>%
  mutate(fish_id = as.character(fish_id)) %>%
  dplyr::select(fish_id, bin, zone, time_in_zone) %>%
  left_join(meta) %>%
  dplyr::filter(Keep == "Y") %>%
  group_by(fish_id, zone) %>%
  mutate(total_timeInZone = sum(time_in_zone)) %>%
  dplyr::distinct(total_timeInZone, .keep_all = TRUE) %>%
  ggplot(aes(x = Genotype, y = total_timeInZone/60)) +
  geom_quasirandom(aes(shape=Sex)
                  ) +
  geom_boxplot(aes(fill = Genotype),
               outlier.shape = NA,
               alpha = 0.5) +
  geom_label_repel(aes(label = fish_id), 
                   fill = NA, 
                   data = . %>% 
                     dplyr::filter(total_timeInZone/60 < 3)) +
  facet_wrap(~zone, nrow = 1) +
  scale_y_log10() +
  ylab("Total time spent in each zone (mins, log scale)") +
  scale_fill_viridis_d() +
  ggtitle("Total time spent in each zone") 


read_csv("data/processed_data/naglu/9m/time_in_zone.csv") %>%
  mutate(fish_id = as.character(fish_id)) %>%
  dplyr::select(fish_id, bin, zone, time_in_zone) %>%
  left_join(meta) %>%
  dplyr::filter(Keep == "Y") %>%
  group_by(fish_id, zone) %>%
  mutate(aveTimeInZone = mean(time_in_zone)) %>%
  dplyr::distinct(aveTimeInZone, .keep_all = TRUE) %>%
  ggplot(aes(x = Genotype, y = aveTimeInZone)) +
  geom_quasirandom(aes(shape=Sex)
                  ) +
  geom_boxplot(aes(fill = Genotype),
               outlier.shape = NA,
                alpha = 0.5) +
  geom_label_repel(aes(label = fish_id), 
                   fill = NA, 
                   data = . %>% 
                     dplyr::filter(aveTimeInZone > 20)) +
  facet_wrap(~zone, nrow = 1) +
  scale_fill_viridis_d() +
  scale_y_log10() +
  ylab("Average time spent in each zone (seconds, log scale") +
  ggtitle("Average time spent in each zone")
```

### model fits

There is no significant effect of genotype on the total or average time spent in each zone. 

```{r}
read_csv("data/processed_data/naglu/9m/time_in_zone.csv") %>%
  mutate(fish_id = as.character(fish_id)) %>%
  dplyr::select(fish_id, bin, zone, time_in_zone) %>%
  left_join(meta) %>%
  dplyr::filter(Keep == "Y") %>%
  group_by(fish_id, zone) %>%
  mutate(total = mean(time_in_zone)) %>%
  dplyr::distinct(total, .keep_all = TRUE) %>%
  lm(log(total) ~ (Genotype + zone)^2 + Sex + Tank,
    data = .) %>%
  summary %>%
  .$coef %>%
  kable(caption = "linear model: log(total) ~ (Genotype + zone)^2 + Sex") %>%
  kable_styling() %>% 
  row_spec(row = c(2,3,9,10), bold = T, color = "black", background = "lightblue")

read_csv("data/processed_data/naglu/9m/time_in_zone.csv") %>%
  mutate(fish_id = as.character(fish_id)) %>%
  dplyr::select(fish_id, bin, zone, time_in_zone) %>%
  left_join(meta) %>%
  dplyr::filter(Keep == "Y") %>%
  group_by(fish_id, zone) %>%
  mutate(aveTimeInZone = mean(time_in_zone)) %>%
  dplyr::distinct(aveTimeInZone, .keep_all = TRUE) %>%
  lm(log(aveTimeInZone) ~ (Genotype + zone)^2 + Sex + Tank,
    data = .) %>%
  summary %>%
  .$coef %>%
  kable(caption = "linear model: log(aveTimeInZone) ~ (Genotype + zone)^2 + Sex") %>%
  kable_styling() %>% 
  row_spec(row = c(2,3,9,10), bold = T, color = "black", background = "lightblue")
```

# check for handedness
[Fontana et al. 2019](https://doi.org/10.1007/s10071-019-01296-9) showed that fish sometimes show a behavioral lateralisation (i.e. handedness). If fish show this, then they would perform less alternation tetragrams not due to working memory.  Here, we defne zebrafish as showing a laterality if they perform left or right turns more than 60% of the total turns.  Indeed, some zebrafish show a preference for L or R turns. However, it is not  significantly to be related to genotype. 

```{r}
# make the LR bias object
LR_Bias <- final_data %>%
  dplyr::select(L, R, total_turns, fish_id) %>%
  group_by(fish_id) %>%
  mutate(L = sum(L),
         R = sum(R),
         total_turns = sum(total_turns),
         L_R_bias = case_when( #consider more than 60% of the time performing a left or right turn to be a bias
           L/total_turns > 0.6 ~ "Left",
           R/total_turns > 0.6 ~ "Right",
           TRUE ~ "Neither"
         )) %>%
  dplyr::select(fish_id, L_R_bias) %>%
  unique() %>%
  mutate(L_R_bias = factor(L_R_bias,
                           levels = c("Neither", "Left", "Right"))
  )

final_data %>%
  left_join(LR_Bias) %>%
  group_by(fish_id) %>%
  mutate(L = sum(L),
         R = sum(R),
         total_turns = sum(total_turns)
  ) %>%
  ggplot(aes(L, R, shape = Genotype)) +
  geom_point(aes(colour = L_R_bias),
             size = 3) +
  geom_abline(slope = 1) +
  scale_color_manual(values = c(viridis_pal(end=0.75)(3), 
                               rep("black", 3))
  )


LR_Bias %>% 
  left_join(meta) %>%
  dplyr::select(Genotype, L_R_bias) %>% 
  group_by(Genotype, L_R_bias) %>% 
  summarise(n = n()) %>% 
  spread(key = "L_R_bias", value = "n") %>% 
  column_to_rownames("Genotype") %>% 
  chisq.test() %>% 
  tidy() %>% 
  mutate(p.value = signif(p.value, 2),
         statistic = signif(statistic, 2)
         ) %>% 
  kable(caption = "Chi-square test of Genotype ~ LR Bias") %>% 
  kable_styling(full_width = F)
```


# test for changes to alternation
Finally, I will now test for alternation changes using a generlatised linear mixed effect model (beta-binomial distribution). We use this because it is count data which is over-dispersed, and fixed and random effects are included. The link function is logit. 

```{r}
glm <-
  final_data %>%
  left_join(LR_Bias) %>%
  mutate(
    dayTime = interaction(Day, Time, drop = T),
    non_alts = total_turns - alts,
    bin = as.factor(bin)
  ) %>%
  glmmTMB(
    cbind(alts, non_alts) ~ (bin + Genotype)^2 + Time + Sex + Tank + L_R_bias + (1|Day) + (1|fish_id) + (1|dayTime),  
    family = betabinomial(),
    data = .
  )

an <- Anova(glm) %>%
  as.data.frame() 

an %>%   
  dplyr::rename(pval = `Pr(>Chisq)`) %>%
  kable() %>%
  kable_styling()
```

# Visualisation of effects
## Genotype

A trend towards improved spatial working memory is observed in the 9m homozygous fish. This is unexpected. 

```{r}
summary(emmeans(glm, specs = "Genotype"), type = "response") %>%
  as_tibble() %>%
  mutate(Genotype = case_when(Genotype == "wt" ~ "+/+",
                              Genotype == "het" ~ "A603fs/+",
                              Genotype == "hom" ~ "A603fs/A603fs")) %>%
  ggplot(aes(Genotype, prob, colour = Genotype)) +
  geom_point(size  = 5,
             position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.25,
                size = 1,
                position = position_dodge(width = 0.5)) +
  labs(y = "Estimated probability of alternation",
       x = "Time interval") +
  theme(axis.text.x = element_text(hjust = 1,
                               vjust = 1,
                               angle = 45)) +
  scale_y_continuous(limits = c(0,0.35)) +
  scale_color_viridis_d(end = 0.8, option = "inferno") +
  ggtitle("GLM predicted probability of zebrafish performing an alternation\ntetragram at 9 months of age",
          subtitle = paste0("Effect of Genotype from GLM: ",
                            signif(an["Genotype","Pr(>Chisq)"], digits = 2)
          )
  )
```
##  Time interval by genotype (interaction)

```{r}
summary(emmeans(glm, specs = "bin", by = "Genotype"), type = "response") %>%
  as_tibble() %>%
  mutate(timeinterval = case_when(
    bin == 1 ~ "0 - 10 mins", 
    bin == 2 ~ "10 - 20 mins", 
    bin == 3 ~ "20 - 30 mins", 
    bin == 4 ~ "30 - 40 mins", 
    bin == 5 ~ "40 - 50 mins", 
    bin == 6 ~ "50 - 60 mins"
  )) %>% 
  ggplot(aes(timeinterval, prob, colour = Genotype, group = Genotype)) +
  geom_point(size = 2,
             position = position_dodge(width = 0.5)) +
  geom_line(position = position_dodge(width = 0.5)) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.5,
    position = position_dodge(width = 0.5)
  ) +
  scale_y_continuous(limits = c(0, 0.4)) +
  ylab("Estimated probability of \nalternation") +
  xlab("Time interval") +
  theme(
    axis.text.x = element_text(hjust = 1,
                               vjust = 1,
                               angle = 45)
  ) + 
  scale_color_viridis_d(end = 0.8, option = "inferno") +
  ggtitle("GLM predicted probability of zebrafish performing an alternation\ntetragram across the 6x10 min bins",
          subtitle = paste0("Effect of bin:Genotype from GLM: ",
                            signif(an["bin:Genotype","Pr(>Chisq)"], digits = 2)
          )
  )
```
## Sex

A trend towards improved spatial working memory is observed in the 9m homozygous fish. This is unexpected. 

```{r}
summary(emmeans(glm, specs = "Sex"), type = "response") %>%
  as_tibble() %>%
  ggplot(aes(Sex, prob, colour = Sex)) +
  geom_point(size  = 5,
             position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.25,
                size = 1,
                position = position_dodge(width = 0.5)) +
  labs(y = "Estimated probability of alternation",
       x = "Time interval") +
  theme(axis.text.x = element_text(hjust = 1,
                               vjust = 1,
                               angle = 45)) +
  scale_y_continuous(limits = c(0,0.35)) +
  scale_color_viridis_d(end = 0.8, option = "inferno") +
  ggtitle("GLM predicted probability of zebrafish performing an alternation\ntetragram at 9 months of age",
          subtitle = paste0("Effect of Sex from GLM: ",
                            signif(an["Sex","Pr(>Chisq)"], digits = 2)
          )
  )
```

## Left/right bias

As previously observed, the effect of fish having a left or right turn bias significantly affects how many alts it does. 

```{r}
summary(emmeans(glm, specs = "L_R_bias"), type = "response") %>%
  as_tibble() %>%
  ggplot(aes(L_R_bias, prob, colour = L_R_bias)) +
  geom_point(size = 5,
             position = position_dodge(width = 0.5)) +
  geom_errorbar(
    aes(ymin = lower.CL, ymax = upper.CL),
    width = 0.25,
    size = 1,
    position = position_dodge(width = 0.25)
    ) +
  ylab("Estimated probability of alternation") +
  xlab("Time interval") +
  theme(
    axis.text.x = element_text(hjust = 1,
                               vjust = 1,
                               angle = 45)
    ) +
  scale_y_continuous(limits = c(0,0.35)) +
  scale_color_viridis_d(end = 0.8, option = "viridis") +
  ggtitle("GLM predicted probability of zebrafish performing an alternation\ntetragram due to having a left or right turn bias",
          subtitle = paste0("Effect of L_R_bias from GLM: ",
                            signif(an["L_R_bias","Pr(>Chisq)"], digits = 2)
          )
  )
```


# Conclusion

9 month old zebrafish homozygous for the *naglu* ^A603fs^ mutation perform more alternations than their wild type siblings. This is indicative of improved spatial working memory. Perhaps we are picking up a hyperactivity phenotype consistent with the observed behaviour of human children living with this disease. 

```{r}
# export data
saveRDS(glm, "data/R objects/naglu9m.rds")
```
